---
title: "MLA CIA1"
author: "Shweta Seth 2027661"
date: "26/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
*Libraries Used*
```{r}
# Important Library used
library(caTools)
library(ggplot2)
library(psych)
library(plyr)
library(descr)
library(caTools)
library(ggplot2)
library(car)
library(lmtest)
library(moments)
library(broom)
library(Metrics)
```
*DATA CLEANING*  
```{r}
# Reading the dataset
Placement_data<-read.csv("C:/Users/User/Desktop/MLA data/Placement_Data_Full_Class.csv",stringsAsFactors = TRUE)
Placement_data
# Dimension of dataset
dim(Placement_data)
# Head values 
head(Placement_data)
# Tail Values
tail(Placement_data)
# Summary Statistics of dataset
summary(Placement_data)
#Replacing null values in Salary column with 0
index<-is.na(Placement_data$salary)
Placement_data$salary[index]<-0
Placement_data$salary

# Head values of dataset
head(Placement_data,10)
# Dimensions of dataset
dim(Placement_data)
#Sum of null values in 'Salary' Column
sum(is.na(Placement_data$salary))
# converting status column into numeric values of 0 and 1 where 1 means Placed , 0 means Not Placed
Placement_data$status = ifelse( Placement_data$status == "Placed", 1, 0)
head(Placement_data)
Placement_data$status
```
*Histogram*
```{r}

# histogram of SSC percentage (12th percentage)
hist(Placement_data$ssc_p)
# histogram of HSC percentage (10th percentage)
hist(Placement_data$hsc_p)
# histogram of Graduation degree percentage 
hist(Placement_data$degree_p)
# histogram of employability test
hist(Placement_data$etest_p)
# histogram of MBA percentage
hist(Placement_data$MBA_per)
# Boxplot of SSC percentage (12th percentage)
boxplot(Placement_data$ssc_p)
# Boxplot of HSC percentage (10th percentage)
boxplot(Placement_data$hsc_p)
# Boxplot of Graduation degree percentage
boxplot(Placement_data$degree_p)
# Boxplot of Salary
boxplot(Placement_data$salary)
```
*Renaming few columns for better understanding*
```{r}
# RENAMING THE COLUMNS
# Renaming 'degree_p' to 'Grad_per'
Placement_data$Grad_per<-Placement_data$degree_p
# Renaming 'degree_t' to 'Grad_stream'
Placement_data$Grad_stream<-Placement_data$degree_t
# Renaming 'ssc_p' to 'SSC_per'
Placement_data$SSC_per<-Placement_data$ssc_p
# Renaming 'SSC_per' to 'ssc_p'
Placement_data$HSC_per<-Placement_data$hsc_p
# Renaming 'hsc_p' to 'HSC_per'
Placement_data$SSC_board<-Placement_data$ssc_b
# Renaming 'hsc_b' to 'HSC_board'
Placement_data$HSC_board<-Placement_data$hsc_b
# Renaming 'status' to 'Placed_status'
Placement_data$Placed_status<-Placement_data$status
# Renaming 'mba_p' to 'MBA_per'
Placement_data$MBA_per<-Placement_data$mba_p

# Summary of dataset
summary(Placement_data)
# View dataset
View(Placement_data)
```
*DATA EXPLORATION* 
```{r}
# CONTINGENCY TABLES
# Table of gender
table(Placement_data$gender)
# Table of SSC_board
table(Placement_data$SSC_board)
# Table of HSC_board
table(Placement_data$HSC_board)
# Table of degree_t
table(Placement_data$degree_t)
# Table of workex
table(Placement_data$workex)
# Table of status
table(Placement_data$status)
# Table of specialisation
table(Placement_data$specialisation)

# Table between gender and SSC_board
table(Placement_data$gender,Placement_data$SSC_board)
# Table between gender and HSC_board
table(Placement_data$gender,Placement_data$HSC_board)
# Table between gender and Placed_status
table(Placement_data$gender,Placement_data$Placed_status)
# Table between gender and specialization
table(Placement_data$gender,Placement_data$specialisation)
# Table between Placed_status and specialization
table(Placement_data$Placed_status,Placement_data$specialisation)
```
*Statistical Tests
```{r}
# STATISTICAL TEST 
# Student T test 
# Salary offered by corporate  is same among males/females
t.test(Placement_data$salary~Placement_data$gender)
#mean in group ( Female ) F 168815.8   and  mean in group (Male) M 215043.2 
t.test(Placement_data$salary~Placement_data$specialisation)
#mean in group Mkt&Fin 236591.7 and  mean in group Mkt&HR 150842.1
 
# Correlation  
cor.test(Placement_data$HSC_per,Placement_data$SSC_per)
# here95 percent confidence interval: 0.4054137 0.6039438
# cor 0.5114721  p- value is less than 0.05 hence we reject H0

cor.test(Placement_data$degree_p,Placement_data$SSC_per)
#95 percent confidence interval:0.4360107 0.6270357   cor 0.538404 
cor.test(Placement_data$degree_p,Placement_data$HSC_per)
#95 percent confidence interval:0.3189315 0.5368211
# cor 0.4342058 

# H0 no association b/w gender and placement status (independent)
# Ha association b/w gender and placement status     (dependent)
chisq.test(Placement_data$status,Placement_data$gender)
# since p-value is greater than 0.05 , we fail to reject H0 , hence no association b/w 2 variables

#NORMALITY CHECK
shapiro.test(Placement_data$Placed_status)
shapiro.test(Placement_data$salary)
shapiro.test(Placement_data$SSC_per)
shapiro.test(Placement_data$HSC_per)
shapiro.test(Placement_data$Grad_per)
shapiro.test(Placement_data$MBA_per)
shapiro.test(Placement_data$etest_p)
```

# VISUALIZATION OF DATASET USING GGPLOT
```{r}
# Bar graph between gender and specialization
ggplot(Placement_data,aes(gender,fill=specialisation))+geom_bar(position = "dodge")
# Bar graph between gender and workex
ggplot(Placement_data,aes(gender,fill=workex))+geom_bar(position = "dodge")
# Bar graph between gender and SSC_board
ggplot(Placement_data,aes(gender,fill=SSC_board))+geom_bar(position = "dodge")
# Bar graph between gender and HSC_board
ggplot(Placement_data,aes(gender,fill=HSC_board))+geom_bar(position = "dodge")
# Bar graph between gender and degree_t
ggplot(Placement_data,aes(gender,fill=degree_t))+geom_bar(position = "dodge")
# Bar graph between gender and hsc_s
ggplot(Placement_data,aes(gender,fill=hsc_s))+geom_bar(position = "dodge")

# Line graph between gender and HSC_per
ggplot(Placement_data, aes(gender,HSC_per))+geom_line()
# Line graph between SSC_board and salary
ggplot(Placement_data, aes(SSC_board, salary))+geom_line()
# Line graph between HSC_board and salary
ggplot(Placement_data, aes(HSC_board, salary))+geom_line()
# Line graph between specialisation and Placed_status
ggplot(Placement_data, aes(specialisation,Placed_status))+geom_line()

# Scatter Plot between HSC_per and SSC_per
ggplot(Placement_data, aes(HSC_per,SSC_per))+geom_point()
ggplot(Placement_data, aes(HSC_per,SSC_per))+geom_point(aes(color=gender))
# Scatter Plot between HSC_per and Grad_per
ggplot(Placement_data, aes(HSC_per,Grad_per))+geom_point(aes(color=gender))
# Scatter Plot between HSC_per and MBA_per
ggplot(Placement_data, aes(HSC_per,MBA_per))+geom_point(aes(color=gender))
# Scatter Plot between SSC_per and MBA_per
ggplot(Placement_data, aes(SSC_per,MBA_per))+geom_point(aes(color=gender))
# Scatter Plot between SSC_per and Grad_per
ggplot(Placement_data, aes(SSC_per,Grad_per))+geom_point(aes(color=gender))
# Scatter Plot between MBA_per and Grad_per
ggplot(Placement_data, aes(MBA_per,Grad_per))+geom_point(aes(color=gender))
# Scatter Plot between etest_p and salary
ggplot(Placement_data, aes(etest_p,salary))+geom_point(aes(color=gender))
```
*MULTIPLE LINEAR REGRESSION*
*mathematical formula for multiple regression*
y = b1x1 + b2x2 + … + bnxn + c.
```{r}
set.seed(100)
# Splitting dataset into 70% training and 30% testing data
split1<-sample.split(Placement_data$sl_no,SplitRatio = 0.7)
summary(split1)
# 70% training data
datatrain<-subset(Placement_data,split1==TRUE)
# 30% testing data
datatest<-subset(Placement_data,split1==FALSE)
# Summary of training data
summary(datatrain)
# Structure of training data
str(datatrain)
# Dimension of training data
dim(datatrain)
# Dimension of testing data
dim(datatest)
#Checking for Linear Relationship using multiple inear regression model
reg_status<-lm(status~ssc_p
         +hsc_p
         +degree_p
         +mba_p
         +etest_p
         +salary,
         data=datatrain)
reg_status
# summary of regression model 
summary(reg_status)

summary(reg_status)$coefficient
# confidence interval of the model coefficient
confint(reg_status)
# Residual Standard Error
sigma(reg_status)/mean(Placement_data$Placed_status)
```
*Multicollinearity*
```{r}
#multicollinearity
vif(reg_status)
# there is no problem of multicollinerity as value of variables is less than 2.5
```
*Randomness of error terms*
```{r}
# checking randomness of error terms
plot(reg_status$residuals,c(1:length(reg_status$residuals)))
```
**
```{r}
#shapiro, qqnorm,boxplot, histogram
qqnorm(reg_status$residuals)
qqline(reg_status$residuals)
boxplot(reg_status$residuals)
#H0 data is normally distributed , Ha data is not normally distributed
shapiro.test(reg_status$residuals)

# normality check through histogram
x<-reg_status$residuals
hist(x)
# dataset is not normally distributed , it is slightly deviated from being normal

#skewness and kurtosis --normality check
datatrain$residual1<-reg_status$residuals
summary(datatrain$residual1)

descr(datatrain)
summary(datatrain)

# Kurtosis and Skewness of dataset
moments::skewness(datatrain$residual1)
moments::kurtosis(datatrain$residual1)
#There is problem of kurtosis and hence we conclude data is not normally distributed
```
*Prediction and RMSE*
```{r}
#prediction and rmse
predicted<-predict(reg_status,Placement_data)
rmse(predicted,Placement_data$Placed_status)
```
*Outliers Detection and Heteroscadicity check*
```{r}
#OUTLIERS
# Checking Heteroscadicity in model 
plot(reg_status$residuals,reg_status$fitted.values)
# if there is cone shape then maybe heterscadicity problem

#H0 equal variance , Ha unequal variance
# problem of normality of error terms and we reject H0
# Breusch Pagan Test 
bptest(reg_status)
```
*Influence Plot*
```{r}
# outliers detection
# Influence plot for understanding whether there are outliers which are affecting the mode
plot(reg_status)
influenceIndexPlot(reg_status)

boxplot.stats(Placement_data$salary)$out
```